Perturbing SU($n$)-symmetric interactions

### 2019-11-01 collective_sun/sun_perturbation.tex
more rearrangements / polishing of the text

We consider an array of $N$ multilevel spins with non-uniform
SU($n$)-symmetric interactions that can be written in the form
\begin{align}
  H_0 = \sum_{p<q} h_{pq} \Pi_{pq}
  = \f12 \sum_{p,q} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
  \label{eq:ints}
\end{align}
where the coefficients $h_{pq}$ in the Hamiltonian are all negative,
with $h_{pq}=h_{qp}$ and $h_{pp}=0$;
$S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
$\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes spins
$p$ and $q$.  The permutation operator $\Pi_{pq}$ is a multilevel
generalization of the SU(2)-symmetric spin interaction
$\m S_p\c\m S_q$.

The ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$
consists of fully symmetric states that are simultaneous $+1$
eigenstates of all permutation operators $\Pi_{pq}$.  We wish to
determine the effective dynamics induced on the fully symmetric
manifold $\M_0$ by weak single- and two-body perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p V_p = \v v\c\v V,
  &&
  \V_2 \equiv \sum_{p<q} w_{pq} X_p Y_q
  = \f12 \sum_{p,q} X_p w_{pq} Y_q
  = \f12 \v X \c\m w \c\v Y,
  \label{eq:perturbations}
\end{align}
where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$;
$\v V,\v X,\v Y$ are respectively vectors of all $V_p,X_p,Y_p$; $\v v$
and $\m w$ are respectively a vector and matrix of all $v_p,w_{pq}$;
we enforce $w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to
denote an inner product between vectors and matrices as appropriate.

We will assume throughout these notes that the interaction Hamiltonian
$H_0$ is gapped, and that the interaction energy difference between
$\M_0$ and its orthogonal complement is much larger than than any
additional couplings that we introduce.  The energy of fully symmetric
states $\ket\psi\in\M_0$ is
\begin{align}
  E_0 = \f12 \sum_{p,q} h_{pq}
  = \f12 \sum_p h_p
  = \f12 N h,
\end{align}
where for later convenience we define the net interaction energy $h_p$
associated with spin $p$, the mean $h$ of all such energies:
\begin{align}
  h_p \equiv \sum_q h_{pq},
  &&
  h \equiv \EE_p\sp{h_p} = \f1N \sum_p h_p.
\end{align}
The effective Hamiltonian $H_M$ induced on the ground-state manifold
$\M_0$ by an $M$-body perturbation $\V_M$ through second order in
perturbation theory is given by\cite{bravyi2011schrieffer,
  perlin2019effective}
\begin{align}
  H_M = H_M^{(1)} + H_M^{(2)},
  &&
  H_M^{(1)} = \P_0 \V_M \P_0,
  &&
  H_M^{(2)} = - \P_0 \V_M \E \V_M \P_0,
  &&
  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace of the
interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
that of fully symmetric manifold $\M_0$; that is, the energy of states
in the image of $\P_\Delta$ is $E_0+\Delta$.

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we expand
\begin{align}
  H_1^{(1)} = \v v \c \P_0 \v V \P_0,
\end{align}
and use the permutational symmetry of the ground-state manifold to
simplify
\begin{align}
  H_1^{(1)} =  \sum_p v_p \P_0 V_0 \P_0 \simeq v V,
  &&
  v \equiv \EE_p\sp{v_p},
  &&
  V \equiv \sum_p V_p,
  \label{eq:H_1_1}
\end{align}
where $\simeq$ denotes equality up to a restriction to the
ground-state manifold $\M_0$.  In order to compute the second order
effective Hamiltonian $H_1^{(2)}$, we first choose an arbitrary fully
symmetric state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \p{\v v\c\m h - \v v\oc\v h}\c \v V \ket\psi,
  \label{eq:H_V1_psi_start}
\end{align}
where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}, and $\v v\oc\v h$ is
the element-wise (Hadamard) product of $\v v$ and $\v h$, i.e.~a
vector with components $v_kh_k$.  We thus find that the vector
$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
if the coefficients $\v v$ satisfy
\begin{align}
  \m h \c \v v - \v h\oc\v v
  = \p{\m h - \diag\v h}\c\v v
  = \Delta\p{\v v} \v v,
  \label{eq:cond_1}
\end{align}
where $\diag\v h$ is a matrix with $\v h$ on the diagonal and zeroes
everywhere else; and $\Delta\p{\v v}$ is some constant that may depend
on $\v v$.  If the interaction Hamiltonian $H_0$ is translationally
invariant, then all $h_q=h$ and $\v h\oc\v v=h\v v$, so the condition
in \eqref{eq:cond_1} simplifies to
\begin{align}
  \m h \c\v v = \sp{\Delta\p{\v v}+h} \v v.
\end{align}
If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta$, then
\begin{align}
  H_0 \V_1\p{\v v_\Delta} \ket\psi
  = \sp{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
  \label{eq:H_V1_psi}
\end{align}
Interestingly, the energy of the (unnormalized) state
$\V_1\p{\v v}\ket\psi$ depends only on the coefficients $\v v$, and is
entirely independent of the state $\ket\psi$ or choice of single-spin
trace-zero operator used to build $\V_1$ in \eqref{eq:perturbations}.
Finding operators $\V_1\p{\v v}$ that generate eigenvectors of the
interaction Hamiltonian $H_0$ when they act on fully symmetric states
$\ket\psi$ thus reduces to finding eigenvectors $\v v$ of the
interaction matrix $\m h-\diag\v h$.

If the interaction Hamiltonian $H_0$ is gapped, then a vector
$\V_1\p{\v v_\Delta}\ket\psi$ with $\Delta=0$ must lie within the
fully symmetric manifold $\M_0$, which implies that the operator
$\V_1\p{\v v_\Delta}$ preserves the permutational symmetry of $\M_0$.
The only way for $\V_1\p{\v v_\Delta}$ to obey permutational symmetry
is for $\v v_\Delta$ to be a constant vector,
i.e.~$\v v_\Delta=v_\Delta\v\1$ with $\1$ a constant vector of ones.
Indeed, a constant vector $\v\1$ satisfies the condition in
\eqref{eq:cond_1} with eigenvalue $0$.  All other vectors
$\v v_\Delta$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must be
orthogonal to $\v\1$, and therefore mean-zero.

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
expanded into its projections $\v v_\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
\begin{align}
  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
\end{align}
where each operator $\V_1\p{\v v_\Delta}$ generates a state with
interaction energy $\Delta$ above that of the fully symmetric
manifold.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = -\P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta}
  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
\end{align}
where the the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out
in Appendix \ref{sec:PVVP}, giving us
\begin{align}
  H_1^{(2)} \simeq \f1{N-1} \sum_{\Delta>0}
  \f{\var\p{\v v_\Delta}}{\Delta} \p{V^2 - N W}.
  \label{eq:H_1_2}
\end{align}
Here $\simeq$ denotes equality up to a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2}
  = \EE_p\sp{x_p^2} - \EE_q\sp{x_q}^2,
  &&
  W \equiv \sum_p V_p^2,
\end{align}
with $\EE_p\sp{x_p}\equiv\sum_px_p/N$ the mean of $\v x$.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} can be computed using the permutational
symmetry of $\M_0$ (see Appendix \ref{sec:PXYP}):
\begin{align}
  H_2^{(1)}
  = \f12 \P_0 \v X\c\m w\c\v Y\P_0
  \simeq \f12 w \p{\tilde Z - Z},
\end{align}
where in terms of the anti-commutator $\sp{A,B}_+\equiv AB+BA$ and
collective operators $X\equiv\sum_p X_p$ and $Y\equiv\sum_p Y_p$, we
define
\begin{align}
  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
  &&
  \tilde Z \equiv \f12 \sp{X,Y}_+
  &&
  Z \equiv \sum_p \f12 \sp{X_p, Y_p}_+.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
  - \v\1 \c \p{\m h\oc\m w} \c \v Z \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where $\v\1$ is an $N$-component vector of ones,
\begin{align}
  \bar{\m h}
  \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h},
\end{align}
is a modified interaction matrix with entries $h_{pq}-\p{h_p+h_q}/2$;
and $\m x\oc\m y$ is an element-wise (Hadamard) product of $\m x$ and
$\m y$, with entries $x_{pq}y_{pq}$.  We thus find that the vector
$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
if the matrix $\m w$ of all $w_{pq}$ satisfies
\begin{align}
  \bar{\m h}\oc\m w + \m h\c\m w = \f12 \Delta\p{\m w} \m w,
  \label{eq:cond_2}
\end{align}
for some constant $\Delta\p{\m w}$ that may depend on $\m w$; $\v\1$
is a constant vector of ones.  Remembering that both $\m h$ and $\m w$
are both symmetric (under transposition) with zeros on the diagonal,
the diagonal components of the condition in \eqref{eq:cond_2}
automatically enforce that the last term in \eqref{eq:H_V2_psi}
vanishes.

We can re-cast the condition in \eqref{eq:cond_2} as an eigenvalue
problem by writing the coefficient matrix $\m w$ as a vector in
$\RR^{N^2}$,
\begin{align}
  \m w = \sum_{p,q} w_{pq} \op{p}{q}
  \to \v{\m w} \equiv \sum_{p,q} w_{pq} \ket{pq},
\end{align}
and defining the matrix
\begin{align}
  \check{\m h}
  \equiv \sum_{p,q} \bar h_{pq} \op{pq}
  + \sum_{p,q,\ell} h_{pq} \op{p\ell}{q\ell},
  \label{eq:h_super_mat}
\end{align}
in terms of which the condition in \eqref{eq:cond_2} becomes the
eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \f12 \Delta\p{\m w} \v{\m w}.
  \label{eq:cond_2_eig}
\end{align}
The matrix $\check{\m h}$ in \eqref{eq:h_super_mat} can be written in
the block-diagonal form
\begin{align}
  \check{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
  &&
  \m h_\ell
  \equiv \sum_{p,q}\p{\delta_{pq} \bar h_{p\ell} + h_{pq}} \op{p}{q},
\end{align}
which reduces the eigenvalue equation in \eqref{eq:cond_2_eig} to the
$N$ smaller, uncoupled eigenvalue equations
\begin{align}
  \m h_\ell \c \v{\m w}_\ell = \f12 \Delta\p{\m w} \v w_\ell,
  &&
  \v w_\ell \equiv \sum_p w_{p\ell} \ket{p}.
  \label{eq:cond_2_eig_block}
\end{align}
We can thus (numerically) diagonalize the blocks $\m h_\ell$
independently, and classify their eigenvectors $\v w_\ell$ by the
corresponding eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors
$\v{\m w}$ of $\check{\m h}$ are then built from eigenvectors
$\v{\m w}_\ell$ within each block $\m h_\ell$ that share the same
eigenvalue $\Delta\p{\m w}/2$.  The task of constructing eigenvectors
$\V_2\p{\m w}\ket\psi$ of $H_0$ and determining their energy thus
reduces to finding the eigenvectors and eigenvalues of $N$ matrices
$\m h_\ell$ of size $N\times N$.

@@

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta/2$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \sp{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi.
  \label{eq:H_V1_psi}
\end{align}
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta/2$,
i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta}.
\end{align}
We can then simplify
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
terms collective operators using techniques similar to those in
Appendix \ref{sec:sym_prod}.

@@

Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify the product
\begin{align}
  \P_0 \V_2 \P_0
  = \f12 \P_0 \v X\c\m w\c\v Y\P_0
  = \f12 \sum_{p\ne q} w_{pq} \P_0 X_p Y_q \P_0
  = {N\choose 2} w \P_0 X_0 Y_1 \P_0,
  \label{eq:PXYP_start}
\end{align}
where ${N\choose2}=N\p{N-1}/2$ and
\begin{align}
  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
\end{align}
is the mean value of the pair-wise coefficients in $\V_2$.  To
simplify the operator content of \eqref{eq:PXYP_start}, we define
\begin{align}
  Z_p \equiv \f12\sp{X_p, Y_p}_+,
  &&
  Z \equiv \sum_p Z_p,
  &&
  \tilde Z \equiv \f12 \sp{X, Y}_+,
\end{align}
and expand
\begin{align}
  \P_0 \tilde Z \P_0 = \P_0 Z \P_0 + N\p{N-1} \P_0 X_0 Y_1 \P_0.
\end{align}
Substituting the product $\P_0 X_0 Y_1 \P_0$ back into
\eqref{eq:PXYP_start}, we thus find that
\begin{align}
  \P_0 \V_2 \P_0
  = \f12 w \P_0 \p{\tilde Z -  Z } \P_0.
\end{align}

>>>>

We consider an array of $N$ multilevel spins with non-uniform
SU($n$)-symmetric interactions that can be written in the form
\begin{align}
  H_0 = \sum_{p<q} h_{pq} \Pi_{pq}
  = \f12 \sum_{p,q} h_{pq} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
  \label{eq:ints}
\end{align}
where the coefficients $h_{pq}$ in the Hamiltonian are all negative,
with $h_{pq}=h_{qp}$ and $h_{pp}=0$;
$S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
$\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes spins
$p$ and $q$.  The permutation operator $\Pi_{pq}$ is a multilevel
generalization of the SU(2)-symmetric spin interaction
$\m S_p\c\m S_q$.

The ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$
consists of fully symmetric states that are simultaneous $+1$
eigenstates of all permutation operators $\Pi_{pq}$.  We wish to
determine the effective dynamics induced on the fully symmetric
manifold $\M_0$ by weak single- and two-body perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p V_p = \v v\c\v V,
  &&
  \V_2 \equiv \sum_{p<q} w_{pq} X_p Y_q
  = \f12 \sum_{p,q} X_p w_{pq} Y_q
  = \f12 \v X \c\m w \c\v Y,
  \label{eq:perturbations}
\end{align}
where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$, with
e.g.~$V_p$ related to $V_q$ by a permutation of spins $p$ and $q$;
$\v V,\v X,\v Y$ are respectively vectors of all $V_p,X_p,Y_p$; $\v v$
and $\m w$ are respectively a vector and matrix of all $v_p,w_{pq}$;
we enforce $w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to
denote an inner product between vectors and matrices as appropriate.

We will assume throughout these notes that the interaction Hamiltonian
$H_0$ is gapped, and that the interaction energy difference between
$\M_0$ and its orthogonal complement is much larger than than any
additional couplings that we introduce.  The energy of fully symmetric
states $\ket\psi\in\M_0$ is
\begin{align}
  E_0 = \sum_{p<q} h_{pq}.
\end{align}
The effective Hamiltonian $H_M$ induced on the ground-state manifold
$\M_0$ by an $M$-body perturbation $\V_M$ through second order in
perturbation theory is given by\cite{bravyi2011schrieffer,
  perlin2019effective}
\begin{align}
  H_M = H_M^{(1)} + H_M^{(2)},
  &&
  H_M^{(1)} = \P_0 \V_M \P_0,
  &&
  H_M^{(2)} = - \P_0 \V_M \E \V_M \P_0,
  &&
  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace of the
interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
that of fully symmetric manifold $\M_0$; that is, the energy of states
in the image of $\P_\Delta$ is $E_0+\Delta$.

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we expand
\begin{align}
  H_1^{(1)} = \v v \c \P_0 \v V \P_0,
\end{align}
and use the permutational symmetry of the ground-state manifold to
simplify
\begin{align}
  H_1^{(1)} =  \sum_p v_p \P_0 V_0 \P_0 \simeq v V,
  &&
  v \equiv \EE_p\sp{v_p} = \f1N \sum_p v_p,
  &&
  V \equiv \sum_p V_p,
  \label{eq:H_1_1}
\end{align}
where $\simeq$ denotes equality up to a restriction to the
ground-state manifold $\M_0$, and $\EE_p\sp{x_p}$ denotes the mean
value of $x_p$.  In order to compute the second order effective
Hamiltonian $H_1^{(2)}$, we first choose an arbitrary fully symmetric
state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \p{\v v\c\m h - \v v\oc\v h}\c \v V \ket\psi,
  \label{eq:H_V1_psi_start}
\end{align}
where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of the net interaction energies $h_p\equiv\sum_qh_{pq}$ associated
with individual spins $p$; and $\v v\oc\v h$ is the element-wise
(Hadamard) product of $\v v$ and $\v h$, i.e.~a vector with components
$v_kh_k$.  We thus find that the vector $\V_1\ket\psi$ is an
eigenvector of the interaction Hamiltonian $H_0$ if the coefficients
$\v v$ satisfy
\begin{align}
  \m h \c \v v - \v h\oc\v v
  = \p{\m h - \diag\v h}\c\v v
  = \Delta\p{\v v} \v v,
  \label{eq:cond_1}
\end{align}
where $\diag\v h$ is a matrix with $\v h$ on the diagonal and zeroes
everywhere else; and $\Delta\p{\v v}$ is some constant that may depend
on $\v v$.  If the interaction Hamiltonian $H_0$ is translationally
invariant, then all $h_p=h\equiv\EE_q\sp{h_q}$ and
$\v h\oc\v v=h\v v$, so the condition in \eqref{eq:cond_1} simplifies
to
\begin{align}
  \m h \c\v v = \sp{\Delta\p{\v v}+h} \v v.
\end{align}
If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta$, then
\begin{align}
  H_0 \V_1\p{\v v_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
  \label{eq:H_V1_psi}
\end{align}
Interestingly, the energy of the (unnormalized) state
$\V_1\p{\v v_\Delta}\ket\psi$ depends only on the coefficients
$\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
choice of single-spin trace-zero operator $V_p$ used to build $\V_1$
in \eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
that generate eigenvectors of the interaction Hamiltonian $H_0$ when
they act on fully symmetric states $\ket\psi$ thus reduces to finding
eigenvectors $\v v_\Delta$ of the interaction matrix $\m h-\diag\v h$.

If the interaction Hamiltonian $H_0$ is gapped, then a vector
$\V_1\p{\v v_\Delta}\ket\psi$ with $\Delta=0$ must lie within the
fully symmetric manifold $\M_0$, which implies that the operator
$\V_1\p{\v v_\Delta}$ preserves the permutational symmetry of $\M_0$.
Indeed, a constant vector $\v\1\equiv\p{1,1,1,\cdots}$ of ones
satisfies the condition in \eqref{eq:cond_1} with eigenvalue $0$.  All
other vectors $\v v_\Delta$ satisfying \eqref{eq:cond_1} with
$\Delta\ne0$ must be orthogonal to $\v\1$, and therefore mean-zero.

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
expanded into its projections $\v v_\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
\begin{align}
  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
\end{align}
where each operator $\V_1\p{\v v_\Delta}$ generates a state with
definite interaction energy $\Delta$ above that of the fully symmetric
manifold.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = -\P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta}
  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PVVP}, giving us
\begin{align}
  H_1^{(2)} \simeq \f{\tilde v}{N-1} \p{V^2 - N W},
  &&
  \tilde v \equiv \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta}.
  \label{eq:H_1_2}
\end{align}
Here $\simeq$ denotes equality up to a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2}
  = \EE_p\sp{x_p^2} - \EE_q\sp{x_q}^2,
  &&
  W \equiv \sum_p V_p^2,
\end{align}
with $\EE_p\sp{x_p}\equiv\sum_px_p/N$ the mean of $\v x$.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} can be computed using the permutational
symmetry of $\M_0$ (see Appendix \ref{sec:PXYP}):
\begin{align}
  H_2^{(1)}
  = \f12 \P_0 \v X\c\m w\c\v Y\P_0
  \simeq \f12 w \p{\tilde Z - Z},
\end{align}
where in terms of the anti-commutator $\sp{A,B}_+\equiv AB+BA$ and
collective operators $X\equiv\sum_p X_p$ and $Y\equiv\sum_p Y_p$, we
define
\begin{align}
  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
  &&
  \tilde Z \equiv \f12 \sp{X,Y}_+,
  &&
  Z \equiv \sum_p \f12 \sp{X_p, Y_p}_+.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
  - \v\1 \c \p{\m h\oc\m w} \c \v Z \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where $\v\1$ is an $N$-component vector of ones;
\begin{align}
  \bar{\m h}
  \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h}
\end{align}
is a modified interaction matrix with entries $h_{pq}-\p{h_p+h_q}/2$;
and $\m x\oc\m y$ is an element-wise (Hadamard) product of $\m x$ and
$\m y$, with entries $x_{pq}y_{pq}$.  We thus find that the vector
$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
if the matrix $\m w$ of all $w_{pq}$ satisfies
\begin{align}
  \bar{\m h}\oc\m w + \m h\c\m w = \f12 \Delta\p{\m w} \m w,
  \label{eq:cond_2}
\end{align}
for some constant $\Delta\p{\m w}$ that may depend on $\m w$.
Remembering that the matrices $\m h$ and $\m w$ are both symmetric
(under transposition) with zeros on the diagonal, the diagonal
components of the condition in \eqref{eq:cond_2} automatically enforce
that the last term ($\sim\v Z$) in \eqref{eq:H_V2_psi} vanishes.

We can recast the condition in \eqref{eq:cond_2} as an eigenvalue
problem by writing the coefficient matrix $\m w$ as a vector in
$\RR^{N^2}$,
\begin{align}
  \m w = \sum_{p,q} w_{pq} \op{p}{q}
  \to \v{\m w} \equiv \sum_{p,q} w_{pq} \ket{pq},
\end{align}
and defining the matrix
\begin{align}
  \check{\m h}
  \equiv \sum_{p,q} \bar h_{pq} \op{pq}
  + \sum_{p,q,\ell} h_{pq} \op{p\ell}{q\ell},
  \label{eq:h_super_mat}
\end{align}
in terms of which the condition in \eqref{eq:cond_2} becomes the
eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \f12 \Delta\p{\m w} \v{\m w}.
  \label{eq:cond_2_eig}
\end{align}
The matrix $\check{\m h}$ in \eqref{eq:h_super_mat} can be written in
the block-diagonal form
\begin{align}
  \check{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
  &&
  \m h_\ell
  \equiv \sum_{p,q}\p{\delta_{pq} \bar h_{p\ell} + h_{pq}} \op{p}{q},
\end{align}
which reduces the eigenvalue equation in \eqref{eq:cond_2_eig} to the
$N$ smaller, independent eigenvalue equations
\begin{align}
  \m h_\ell \c \v{\m w}_\ell = \f12 \Delta\p{\m w} \v w_\ell,
  &&
  \v w_\ell \equiv \sum_p w_{p\ell} \ket{p}.
  \label{eq:cond_2_eig_block}
\end{align}
We can thus (numerically) diagonalize each of the blocks $\m h_\ell$,
and classify their eigenvectors $\v w_\ell$ by the corresponding
eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors $\v{\m w}$ of
$\check{\m h}$ are then built from eigenvectors $\v w_\ell$ within
each block $\m h_\ell$ that share the same eigenvalue
$\Delta\p{\m w}/2$.  The task of constructing eigenvectors
$\V_2\p{\m w}\ket\psi$ of $H_0$ and determining their energy thus
reduces to finding the eigenvectors and eigenvalues of $N$ matrices
$\m h_\ell$ of size $N\times N$.

@@

If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2} or
\eqref{eq:cond_2_eig} with eigenvalue $\Delta/2$, then
\begin{align}
  H_0 \V_2\p{\m w_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
  \label{eq:H_V1_psi}
\end{align}
which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
definite energy that depends only on the coefficients $\m w_\Delta$.
As with the case of single-body perturbations, we can decompose any
matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta/2$,
i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
\begin{align}
  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
\end{align}
and in turn simplify
\begin{align}
  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0.
\end{align}
The product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in terms
collective operators using techniques similar to those in Appendix
\ref{sec:sym_prod}, which should yield a effective Hamiltonian of the
form
\begin{align}
  H_2^{(2)} \simeq f\p{\m w_\Delta} \O\p{\v X,\v Y},
\end{align}
where $f\p{\m w_\Delta}$ is a scalar and $\O\p{\v X,\v Y}$ is a
collective operator that is independent of $\m w_\Delta$.

@@

Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify the product
\begin{align}
  \P_0 \V_2 \P_0
  = \f12 \sum_{p\ne q} w_{pq} \P_0 X_p Y_q \P_0
  = {N\choose 2} w \P_0 X_0 Y_1 \P_0,
  \label{eq:PXYP_start}
\end{align}
where ${N\choose2}=N\p{N-1}/2$ and
\begin{align}
  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
\end{align}
is the mean value of the pair-wise coefficients in $\V_2$.  To
simplify the operator content of \eqref{eq:PXYP_start}, we define
\begin{align}
  X \equiv \sum_p X_p,
  &&
  Y \equiv \sum_p Y_p,
  &&
  Z_p \equiv \f12\sp{X_p, Y_p}_+,
  &&
  Z \equiv \sum_p Z_p,
  &&
  \tilde Z \equiv \f12 \sp{X, Y}_+,
\end{align}
and expand
\begin{align}
  \P_0 \tilde Z \P_0 = \P_0 Z \P_0 + N\p{N-1} \P_0 X_0 Y_1 \P_0.
\end{align}
Substituting the product $\P_0 X_0 Y_1 \P_0$ back into
\eqref{eq:PXYP_start}, we thus find that
\begin{align}
  \P_0 \V_2 \P_0 = \f12 w \P_0 \p{\tilde Z -  Z } \P_0.
\end{align}
